---
title: "kw and boxplots"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data
```{r, echo = FALSE, warning = FALSE}
library(tidyverse)
setwd("~/Documents")
DOMINO <- read_csv(file = "FFQ_for_MET582.csv", col_names = TRUE, na = "NA") %>%
  dplyr::select(COUNTRY:OLIVES)

DOMINO[DOMINO == -9] <- NA
DOMINO[DOMINO == -8] <- NA
DOMINO[DOMINO == 1] <- 0
DOMINO[DOMINO == 2] <- 0.067
DOMINO[DOMINO == 3] <- 0.143
DOMINO[DOMINO == 4] <- 0.429
DOMINO[DOMINO == 5] <- 0.756
DOMINO[DOMINO == 6] <- 1
DOMINO[DOMINO == 7] <- 2.5
DOMINO[DOMINO == 8] <- 4.5
DOMINO[DOMINO == 9] <- 6

DOMINO <- rowid_to_column(DOMINO, "PARTICIPANT")
```

## Add food group category to DOMINO
```{r, echo = FALSE}
#### Group by major food groups by adding category column to DOMINO table
MF <- colnames(DOMINO)[3:29]
BSB <- colnames(DOMINO)[30:36]
C <- colnames(DOMINO)[37:40]
PRP <- colnames(DOMINO)[41:58]
DF <- colnames(DOMINO)[59:80]
OBV <- colnames(DOMINO)[81:89]
SwS <- colnames(DOMINO)[90:116]
SSS <- colnames(DOMINO)[117:129]
D <- colnames(DOMINO)[130:147]
Fr <- colnames(DOMINO)[148:164]
V <- colnames(DOMINO)[165:197]

DOMINO_category <- pivot_longer(DOMINO, BEEF:OLIVES, names_to = "FOOD", values_to = "daily_FFQ_score") %>%
  mutate(category = case_when(FOOD %in% MF ~ "meat_fish",
                            FOOD %in% BSB ~ "bread_sav_biscuits",
                            FOOD %in% C ~ "cereals",
                            FOOD %in% PRP ~ "potatoes_rice_pasta",
                            FOOD %in% DF ~ "dairy_fats",
                            FOOD %in% OBV ~ "on_bread_veg",
                            FOOD %in% SwS ~ "sweets_snacks",
                            FOOD %in% SSS ~ "soups_sauces_spreads",
                            FOOD %in% D ~ "drinks",
                            FOOD %in% Fr ~ "fruits",
                            FOOD %in% V ~ "vegetables")) 
```

## Test data normality
```{r}
library(ggplot2)
ggplot(data = DOMINO_category) +
  geom_histogram(aes(x = daily_FFQ_score), binwidth = 1) +
  theme_bw()

qqnorm(DOMINO_category$daily_FFQ_score, pch = 1, frame = FALSE)
```

Anderson-Darling test for normality used as DOMINO_category daily data contains more than 5000  values.

H0: data are normally distributed

H1: data are not normally distributed
```{r}
library(nortest)
ad.test(DOMINO_category$daily_FFQ_score)
```
p-value < 0.05, therefore accept the alternate hypothesis.

All three of these measures of normality indicate that the data are not normally distributed so non-parametric statistical tests should be used for analyses.

## Kruskal-Wallis tests within groups

Kruskal-Wallis tests used as the data is non-parametric and ordinal.

H0: median food intake does not differ between countries for each food category

H1: median food intake does differ between countries for each food category

```{r, echo = FALSE}
DOMINO_MF <- filter(DOMINO_category, category == "meat_fish")
KW_MF <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_MF)

DOMINO_BSB <- filter(DOMINO_category, category == "bread_sav_biscuits")
KW_BSB <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_BSB)

DOMINO_C <- filter(DOMINO_category, category == "cereals")
KW_C <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_C)

DOMINO_PRP <- filter(DOMINO_category, category == "potatoes_rice_pasta")
KW_PRP <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_PRP)

DOMINO_DF <- filter(DOMINO_category, category == "dairy_fats")
KW_DF <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_DF)

DOMINO_OBV <- filter(DOMINO_category, category == "on_bread_veg")
KW_OBV <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_OBV)

DOMINO_SwS <- filter(DOMINO_category, category == "sweets_snacks")
KW_SwS <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_SwS)

DOMINO_SSS <- filter(DOMINO_category, category == "soups_sauces_spreads")
KW_SSS <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_SSS)

DOMINO_D <- filter(DOMINO_category, category == "drinks")
KW_D <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_D)

DOMINO_Fr <- filter(DOMINO_category, category == "fruits")
KW_Fr <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_Fr)

DOMINO_V <- filter(DOMINO_category, category == "vegetables")
KW_V <- kruskal.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_V)

KW_p_value <- c(KW_MF$p.value, KW_BSB$p.value, KW_C$p.value, KW_PRP$p.value, KW_DF$p.value,
          KW_OBV$p.value, KW_SwS$p.value, KW_SSS$p.value, KW_D$p.value, KW_Fr$p.value,
          KW_V$p.value)
Food_Categories <- c("meat_fish", "bread_sav_biscuits", "cereals", "potatoes_rice_pasta",
                     "dairy_fats", "on_bread_veg", "sweets_snacks", "soups_sauces_spreads",
                     "drinks", "fruits", "vegetables")
KW_p_category <- tibble(Food_Categories, KW_p_value)
KW_p_category
```

## Boxplots
```{r}
foodgroup_boxplot <- function(cat) {
  DOMINO_category %>%
    filter(category == cat) %>%
    ggplot(aes(x = COUNTRY, y = daily_FFQ_score, colour = COUNTRY)) +
    geom_boxplot() +
    ylim(0, 6) +
    ylab("Daily intake derived from FFQ score") +
    theme_bw() 
}
print(foodgroup_boxplot("meat_fish") + labs(title= "Meat and fish intake by country"))
print(foodgroup_boxplot("bread_sav_biscuits") + labs(title= "Bread and savoury biscuits intake by country"))
print(foodgroup_boxplot("cereals") + labs(title= "Cereals intake by country"))
print(foodgroup_boxplot("potatoes_rice_pasta") + labs(title= "Potatoes, rice and pasta intake by country"))
print(foodgroup_boxplot("dairy_fats") + labs(title= "Dairy and fat intake by country"))
print(foodgroup_boxplot("on_bread_veg") + labs(title= "Foods on bread and vegetables intake by country"))
print(foodgroup_boxplot("sweets_snacks") + labs(title= "Sweet and snack intake by country"))
print(foodgroup_boxplot("soups_sauces_spreads") + labs(title= "Soups, sauces and spreads intake by country"))
print(foodgroup_boxplot("drinks") + labs(title= "Drink intake by country"))
print(foodgroup_boxplot("fruits") + labs(title= "Fruit intake by country"))
print(foodgroup_boxplot("vegetables") + labs(title= "Vegetable intake by country"))
```

## Transformation

The boxplots demonstrate that the data is right-skewed so a logarithmic transformation of the daily FFQ score + 1 (so data points of 0 can be logarithmically transformed) was used to aid data visualisation.

```{r}
DOMINO_transformation <- mutate(DOMINO_category, trans_daily_FFQ_score = log(daily_FFQ_score + 1))
foodgroup_boxplot <- function(cat) {
  DOMINO_transformation %>%
    filter(category == cat) %>%
    ggplot(aes(x = COUNTRY, y = trans_daily_FFQ_score, colour = COUNTRY)) +
    geom_boxplot() +
    ylim(0, 2) +
    ylab("Transformed daily intake derived from FFQ score") +
    theme_bw() 
}

print(foodgroup_boxplot("meat_fish") + labs(title= "Meat and fish intake by country"))
print(foodgroup_boxplot("bread_sav_biscuits") + labs(title= "Bread and savoury biscuits intake by country"))
print(foodgroup_boxplot("cereals") + labs(title= "Cereals intake by country"))
print(foodgroup_boxplot("potatoes_rice_pasta") + labs(title= "Potatoes, rice and pasta intake by country"))
print(foodgroup_boxplot("dairy_fats") + labs(title= "Dairy and fat intake by country"))
print(foodgroup_boxplot("on_bread_veg") + labs(title= "Foods on bread and vegetables intake by country"))
print(foodgroup_boxplot("sweets_snacks") + labs(title= "Sweet and snack intake by country"))
print(foodgroup_boxplot("soups_sauces_spreads") + labs(title= "Soups, sauces and spreads intake by country"))
print(foodgroup_boxplot("drinks") + labs(title= "Drink intake by country"))
print(foodgroup_boxplot("fruits") + labs(title= "Fruit intake by country"))
print(foodgroup_boxplot("vegetables") + labs(title= "Vegetable intake by country"))
```

## Pairwise comparisons
Using Wilcoxon rank sum test (Mann-Whitney U test) as the distribution is non-parametric and the
data aren't paired.

H0: median daily food intake does not differ between (2 selected) countries for each category

H1: median daily food intake does differ between (2 selected) countries for each category

### bread and savoury biscuits
```{r}
wilcox_pairwise <- function(cat, country1, country2) {
  DOMINO_category2 <- filter(DOMINO_category, category == cat)
  DOMINO_category2 <- filter(DOMINO_category2, COUNTRY == country1 | COUNTRY == country2)
  wilcox.test(daily_FFQ_score ~ COUNTRY, data = DOMINO_category2)
}

UK_Sw_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "UK", country2 = "Switzerland")
UK_Po_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "UK", country2 = "Poland")
UK_Sp_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "UK", country2 = "Spain")
Sw_Po_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "Switzerland", country2 = "Spain")
Sp_Po_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "Spain", country2 = "Poland")
wilcox_BSB <- c(UK_Sw_BSB_w$p.value, UK_Po_BSB_w$p.value, UK_Sp_BSB_w$p.value, 
                Sw_Po_BSB_w$p.value, Sw_Sp_BSB_w$p.value, Sp_Po_BSB_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_BSB_p <- tibble(Country1, Country2, wilcox_BSB)
wilcox_BSB_p$holm_adjusted_w_BSB <- p.adjust(wilcox_BSB, method = "holm")
wilcox_BSB_p
```
### cereals
```{r}
UK_Sw_C_w <- wilcox_pairwise(cat = "cereals", country1 = "UK", country2 = "Switzerland")
UK_Po_C_w <- wilcox_pairwise(cat = "cereals", country1 = "UK", country2 = "Poland")
UK_Sp_C_w <- wilcox_pairwise(cat = "cereals", country1 = "UK", country2 = "Spain")
Sw_Po_C_w <- wilcox_pairwise(cat = "cereals", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_C_w <- wilcox_pairwise(cat = "cereals", country1 = "Switzerland", country2 = "Spain")
Sp_Po_C_w <- wilcox_pairwise(cat = "cereals", country1 = "Spain", country2 = "Poland")
wilcox_C <- c(UK_Sw_C_w$p.value, UK_Po_C_w$p.value, UK_Sp_C_w$p.value, 
                Sw_Po_C_w$p.value, Sw_Sp_C_w$p.value, Sp_Po_C_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_C_p <- tibble(Country1, Country2, wilcox_C)
wilcox_C_p$holm_adjusted_w_C <- p.adjust(wilcox_C, method = "holm")
wilcox_C_p
```
### potatoes, rice and pasta
```{r}
UK_Sw_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "UK", country2 = "Switzerland")
UK_Po_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "UK", country2 = "Poland")
UK_Sp_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "UK", country2 = "Spain")
Sw_Po_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "Switzerland", country2 = "Spain")
Sp_Po_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "Spain", country2 = "Poland")
wilcox_PRP <- c(UK_Sw_PRP_w$p.value, UK_Po_PRP_w$p.value, UK_Sp_PRP_w$p.value, 
                Sw_Po_PRP_w$p.value, Sw_Sp_PRP_w$p.value, Sp_Po_PRP_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_PRP_p <- tibble(Country1, Country2, wilcox_PRP)
wilcox_PRP_p$holm_adjusted_w_PRP <- p.adjust(wilcox_PRP, method = "holm")
wilcox_PRP_p
```
### dairy and fats
```{r}
UK_Sw_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "UK", country2 = "Switzerland")
UK_Po_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "UK", country2 = "Poland")
UK_Sp_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "UK", country2 = "Spain")
Sw_Po_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "Switzerland", country2 = "Spain")
Sp_Po_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "Spain", country2 = "Poland")
wilcox_DF <- c(UK_Sw_DF_w$p.value, UK_Po_DF_w$p.value, UK_Sp_DF_w$p.value, 
                Sw_Po_DF_w$p.value, Sw_Sp_DF_w$p.value, Sp_Po_DF_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_DF_p <- tibble(Country1, Country2, wilcox_DF)
wilcox_DF_p$holm_adjusted_w_DF <- p.adjust(wilcox_DF, method = "holm")
wilcox_DF_p
```
### sweets and snacks
```{r}
UK_Sw_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "UK", country2 = "Switzerland")
UK_Po_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "UK", country2 = "Poland")
UK_Sp_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "UK", country2 = "Spain")
Sw_Po_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "Switzerland", country2 = "Spain")
Sp_Po_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "Spain", country2 = "Poland")
wilcox_SwS <- c(UK_Sw_SwS_w$p.value, UK_Po_SwS_w$p.value, UK_Sp_SwS_w$p.value, 
                Sw_Po_SwS_w$p.value, Sw_Sp_SwS_w$p.value, Sp_Po_SwS_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_SwS_p <- tibble(Country1, Country2, wilcox_SwS)
wilcox_SwS_p$holm_adjusted_w_SwS <- p.adjust(wilcox_SwS, method = "holm")
wilcox_SwS_p
```
### soups, sauces and spreads
```{r}
UK_Sw_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "UK", country2 = "Switzerland")
UK_Po_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "UK", country2 = "Poland")
UK_Sp_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "UK", country2 = "Spain")
Sw_Po_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "Switzerland", country2 = "Spain")
Sp_Po_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "Spain", country2 = "Poland")
wilcox_SSS <- c(UK_Sw_SSS_w$p.value, UK_Po_SSS_w$p.value, UK_Sp_SSS_w$p.value, 
                Sw_Po_SSS_w$p.value, Sw_Sp_SSS_w$p.value, Sp_Po_SSS_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_SSS_p <- tibble(Country1, Country2, wilcox_SSS)
wilcox_SSS_p$holm_adjusted_w_SSS <- p.adjust(wilcox_SSS, method = "holm")
wilcox_SSS_p
```
### drinks
```{r}
UK_Sw_D_w <- wilcox_pairwise(cat = "drinks", country1 = "UK", country2 = "Switzerland")
UK_Po_D_w <- wilcox_pairwise(cat = "drinks", country1 = "UK", country2 = "Poland")
UK_Sp_D_w <- wilcox_pairwise(cat = "drinks", country1 = "UK", country2 = "Spain")
Sw_Po_D_w <- wilcox_pairwise(cat = "drinks", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_D_w <- wilcox_pairwise(cat = "drinks", country1 = "Switzerland", country2 = "Spain")
Sp_Po_D_w <- wilcox_pairwise(cat = "drinks", country1 = "Spain", country2 = "Poland")
wilcox_D <- c(UK_Sw_D_w$p.value, UK_Po_D_w$p.value, UK_Sp_D_w$p.value, 
                Sw_Po_D_w$p.value, Sw_Sp_D_w$p.value, Sp_Po_D_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_D_p <- tibble(Country1, Country2, wilcox_D)
wilcox_D_p$holm_adjusted_w_D <- p.adjust(wilcox_D, method = "holm")
wilcox_D_p
```
### fruits
```{r}
UK_Sw_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "UK", country2 = "Switzerland")
UK_Po_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "UK", country2 = "Poland")
UK_Sp_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "UK", country2 = "Spain")
Sw_Po_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "Switzerland", country2 = "Spain")
Sp_Po_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "Spain", country2 = "Poland")
wilcox_Fr <- c(UK_Sw_Fr_w$p.value, UK_Po_Fr_w$p.value, UK_Sp_Fr_w$p.value, 
                Sw_Po_Fr_w$p.value, Sw_Sp_Fr_w$p.value, Sp_Po_Fr_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_Fr_p <- tibble(Country1, Country2, wilcox_Fr)
wilcox_Fr_p$holm_adjusted_w_Fr <- p.adjust(wilcox_Fr, method = "holm")
wilcox_Fr_p
```
### vegetables
```{r}
UK_Sw_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "UK", country2 = "Switzerland")
UK_Po_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "UK", country2 = "Poland")
UK_Sp_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "UK", country2 = "Spain")
Sw_Po_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "Switzerland", country2 = "Spain")
Sp_Po_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "Spain", country2 = "Poland")
wilcox_V <- c(UK_Sw_V_w$p.value, UK_Po_V_w$p.value, UK_Sp_V_w$p.value, 
                Sw_Po_V_w$p.value, Sw_Sp_V_w$p.value, Sp_Po_V_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_V_p <- tibble(Country1, Country2, wilcox_V)
wilcox_V_p$holm_adjusted_w_V <- p.adjust(wilcox_V, method = "holm")
wilcox_V_p
```
