---
title: "median_comparisons"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data
```{r, echo = FALSE, warning = FALSE}
library(tidyverse)
setwd("~/Documents")
DOMINO <- read_csv(file = "FFQ_for_MET582.csv", col_names = TRUE, na = "NA") %>%
  select(COUNTRY:OLIVES)
DOMINO[DOMINO == -9] <- NA
DOMINO[DOMINO == -8] <- NA
```

## Add food group category to DOMINO
```{r, echo = FALSE}
#### Group by major food groups by adding category column to DOMINO table
MF <- colnames(DOMINO)[2:28]
BSB <- colnames(DOMINO)[29:35]
C <- colnames(DOMINO)[36:39]
PRP <- colnames(DOMINO)[40:57]
DF <- colnames(DOMINO)[58:79]
OBV <- colnames(DOMINO)[80:88]
SwS <- colnames(DOMINO)[89:115]
SSS <- colnames(DOMINO)[116:128]
D <- colnames(DOMINO)[129:146]
Fr <- colnames(DOMINO)[147:163]
V <- colnames(DOMINO)[164:196]

DOMINO_category <- pivot_longer(DOMINO, -COUNTRY, names_to = "FOOD", values_to = "FFQ") %>%
mutate(category = case_when(FOOD %in% MF ~ "meat_fish",
                            FOOD %in% BSB ~ "bread_sav_biscuits",
                            FOOD %in% C ~ "cereals",
                            FOOD %in% PRP ~ "potatoes_rice_pasta",
                            FOOD %in% DF ~ "dairy_fats",
                            FOOD %in% OBV ~ "on_bread_veg",
                            FOOD %in% SwS ~ "sweets_snacks",
                            FOOD %in% SSS ~ "soups_sauces_spreads",
                            FOOD %in% D ~ "drinks",
                            FOOD %in% Fr ~ "fruits",
                            FOOD %in% V ~ "vegetables"))
```

## Box plots of each food group
```{r}
foodgroup_boxplot <- function(cat) {
  DOMINO_category %>%
    filter(category == cat) %>%
    ggplot(aes(x = COUNTRY, y = FFQ, colour = COUNTRY)) +
    geom_boxplot() +
    theme_bw() 
}
print(foodgroup_boxplot("meat_fish") + labs(title= "Meat and fish intake by country"))
print(foodgroup_boxplot("bread_sav_biscuits") + labs(title= "Bread and savoury biscuits intake by country"))
print(foodgroup_boxplot("cereals") + labs(title= "Cereals intake by country"))
print(foodgroup_boxplot("potatoes_rice_pasta") + labs(title= "Potatoes, rice and pasta intake by country"))
print(foodgroup_boxplot("dairy_fats") + labs(title= "Dairy and fat intake by country"))
print(foodgroup_boxplot("on_bread_veg") + labs(title= "Foods on bread and vegetables intake by country"))
print(foodgroup_boxplot("sweets_snacks") + labs(title= "Sweet and snack intake by country"))
print(foodgroup_boxplot("soups_sauces_spreads") + labs(title= "Soups, sauces and spreads intake by country"))
print(foodgroup_boxplot("drinks") + labs(title= "Drink intake by country"))
print(foodgroup_boxplot("fruits") + labs(title= "Fruit intake by country"))
print(foodgroup_boxplot("vegetables") + labs(title= "Vegetable intake by country"))
```

```{r}

```

## Kruskal-Wallis tests within groups
```{r, echo = FALSE}
DOMINO_MF <- filter(DOMINO_category, category == "meat_fish")
KW_MF <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_MF)

DOMINO_BSB <- filter(DOMINO_category, category == "bread_sav_biscuits")
KW_BSB <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_BSB)

DOMINO_C <- filter(DOMINO_category, category == "cereals")
KW_C <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_C)

DOMINO_PRP <- filter(DOMINO_category, category == "potatoes_rice_pasta")
KW_PRP <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_PRP)

DOMINO_DF <- filter(DOMINO_category, category == "dairy_fats")
KW_DF <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_DF)

DOMINO_OBV <- filter(DOMINO_category, category == "on_bread_veg")
KW_OBV <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_OBV)

DOMINO_SwS <- filter(DOMINO_category, category == "sweets_snacks")
KW_SwS <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_SwS)

DOMINO_SSS <- filter(DOMINO_category, category == "soups_sauces_spreads")
KW_SSS <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_SSS)

DOMINO_D <- filter(DOMINO_category, category == "drinks")
KW_D <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_D)

DOMINO_Fr <- filter(DOMINO_category, category == "fruits")
KW_Fr <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_Fr)

DOMINO_V <- filter(DOMINO_category, category == "vegetables")
KW_V <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_V)

KW_p <- c(KW_MF$p.value, KW_BSB$p.value, KW_C$p.value, KW_PRP$p.value, KW_DF$p.value,
          KW_OBV$p.value, KW_SwS$p.value, KW_SSS$p.value, KW_D$p.value, KW_Fr$p.value,
          KW_V$p.value)
food_categories <- c("meat_fish", "bread_sav_biscuits", "cereals", "potatoes_rice_pasta",
                     "dairy_fats", "on_bread_veg", "sweets_snacks", "soups_sauces_spreads",
                     "drinks", "fruits", "vegetables")
KW_p_category <- tibble(food_categories, KW_p)
KW_p_category
```

## Pairwise comparisons
### bread and savoury biscuits
```{r}
wilcox_pairwise <- function(cat, country1, country2) {
  DOMINO_category2 <- filter(DOMINO_category, category == cat)
  DOMINO_category2 <- filter(DOMINO_category2, COUNTRY == country1 | COUNTRY == country2)
  wilcox.test(FFQ ~ COUNTRY, data = DOMINO_category2)
}

UK_Sw_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "UK", country2 = "Switzerland")
UK_Po_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "UK", country2 = "Poland")
UK_Sp_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "UK", country2 = "Spain")
Sw_Po_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "Switzerland", country2 = "Spain")
Sp_Po_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "Spain", country2 = "Poland")
wilcox_BSB <- c(UK_Sw_BSB_w$p.value, UK_Po_BSB_w$p.value, UK_Sp_BSB_w$p.value, 
                Sw_Po_BSB_w$p.value, Sw_Sp_BSB_w$p.value, Sp_Po_BSB_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_BSB_p <- tibble(Country1, Country2, wilcox_BSB)
wilcox_BSB_p$holm_adjusted_w_BSB <- p.adjust(wilcox_BSB, method = "holm")
wilcox_BSB_p
```

### cereals
```{r}
UK_Sw_C_w <- wilcox_pairwise(cat = "cereals", country1 = "UK", country2 = "Switzerland")
UK_Po_C_w <- wilcox_pairwise(cat = "cereals", country1 = "UK", country2 = "Poland")
UK_Sp_C_w <- wilcox_pairwise(cat = "cereals", country1 = "UK", country2 = "Spain")
Sw_Po_C_w <- wilcox_pairwise(cat = "cereals", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_C_w <- wilcox_pairwise(cat = "cereals", country1 = "Switzerland", country2 = "Spain")
Sp_Po_C_w <- wilcox_pairwise(cat = "cereals", country1 = "Spain", country2 = "Poland")
wilcox_C <- c(UK_Sw_C_w$p.value, UK_Po_C_w$p.value, UK_Sp_C_w$p.value, 
                Sw_Po_C_w$p.value, Sw_Sp_C_w$p.value, Sp_Po_C_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_C_p <- tibble(Country1, Country2, wilcox_C)
wilcox_C_p$holm_adjusted_w_C <- p.adjust(wilcox_C, method = "holm")
wilcox_C_p
```

### potatoes, rice and pasta
```{r}
UK_Sw_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "UK", country2 = "Switzerland")
UK_Po_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "UK", country2 = "Poland")
UK_Sp_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "UK", country2 = "Spain")
Sw_Po_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "Switzerland", country2 = "Spain")
Sp_Po_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "Spain", country2 = "Poland")
wilcox_PRP <- c(UK_Sw_PRP_w$p.value, UK_Po_PRP_w$p.value, UK_Sp_PRP_w$p.value, 
                Sw_Po_PRP_w$p.value, Sw_Sp_PRP_w$p.value, Sp_Po_PRP_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_PRP_p <- tibble(Country1, Country2, wilcox_PRP)
wilcox_PRP_p$holm_adjusted_w_PRP <- p.adjust(wilcox_PRP, method = "holm")
wilcox_PRP_p
```
### dairy and fats
```{r}
UK_Sw_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "UK", country2 = "Switzerland")
UK_Po_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "UK", country2 = "Poland")
UK_Sp_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "UK", country2 = "Spain")
Sw_Po_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "Switzerland", country2 = "Spain")
Sp_Po_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "Spain", country2 = "Poland")
wilcox_DF <- c(UK_Sw_DF_w$p.value, UK_Po_DF_w$p.value, UK_Sp_DF_w$p.value, 
                Sw_Po_DF_w$p.value, Sw_Sp_DF_w$p.value, Sp_Po_DF_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_DF_p <- tibble(Country1, Country2, wilcox_DF)
wilcox_DF_p$holm_adjusted_w_DF <- p.adjust(wilcox_DF, method = "holm")
wilcox_DF_p
```
### sweets and snacks
```{r}
UK_Sw_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "UK", country2 = "Switzerland")
UK_Po_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "UK", country2 = "Poland")
UK_Sp_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "UK", country2 = "Spain")
Sw_Po_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "Switzerland", country2 = "Spain")
Sp_Po_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "Spain", country2 = "Poland")
wilcox_SwS <- c(UK_Sw_SwS_w$p.value, UK_Po_SwS_w$p.value, UK_Sp_SwS_w$p.value, 
                Sw_Po_SwS_w$p.value, Sw_Sp_SwS_w$p.value, Sp_Po_SwS_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_SwS_p <- tibble(Country1, Country2, wilcox_SwS)
wilcox_SwS_p$holm_adjusted_w_SwS <- p.adjust(wilcox_SwS, method = "holm")
wilcox_SwS_p
```
### soups, sauces and spreads
```{r}
UK_Sw_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "UK", country2 = "Switzerland")
UK_Po_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "UK", country2 = "Poland")
UK_Sp_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "UK", country2 = "Spain")
Sw_Po_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "Switzerland", country2 = "Spain")
Sp_Po_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "Spain", country2 = "Poland")
wilcox_SSS <- c(UK_Sw_SSS_w$p.value, UK_Po_SSS_w$p.value, UK_Sp_SSS_w$p.value, 
                Sw_Po_SSS_w$p.value, Sw_Sp_SSS_w$p.value, Sp_Po_SSS_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_SSS_p <- tibble(Country1, Country2, wilcox_SSS)
wilcox_SSS_p$holm_adjusted_w_SSS <- p.adjust(wilcox_SSS, method = "holm")
wilcox_SSS_p
```
### drinks
```{r}
UK_Sw_D_w <- wilcox_pairwise(cat = "drinks", country1 = "UK", country2 = "Switzerland")
UK_Po_D_w <- wilcox_pairwise(cat = "drinks", country1 = "UK", country2 = "Poland")
UK_Sp_D_w <- wilcox_pairwise(cat = "drinks", country1 = "UK", country2 = "Spain")
Sw_Po_D_w <- wilcox_pairwise(cat = "drinks", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_D_w <- wilcox_pairwise(cat = "drinks", country1 = "Switzerland", country2 = "Spain")
Sp_Po_D_w <- wilcox_pairwise(cat = "drinks", country1 = "Spain", country2 = "Poland")
wilcox_D <- c(UK_Sw_D_w$p.value, UK_Po_D_w$p.value, UK_Sp_D_w$p.value, 
                Sw_Po_D_w$p.value, Sw_Sp_D_w$p.value, Sp_Po_D_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_D_p <- tibble(Country1, Country2, wilcox_D)
wilcox_D_p$holm_adjusted_w_D <- p.adjust(wilcox_D, method = "holm")
wilcox_D_p
```
### fruits
```{r}
UK_Sw_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "UK", country2 = "Switzerland")
UK_Po_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "UK", country2 = "Poland")
UK_Sp_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "UK", country2 = "Spain")
Sw_Po_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "Switzerland", country2 = "Spain")
Sp_Po_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "Spain", country2 = "Poland")
wilcox_Fr <- c(UK_Sw_Fr_w$p.value, UK_Po_Fr_w$p.value, UK_Sp_Fr_w$p.value, 
                Sw_Po_Fr_w$p.value, Sw_Sp_Fr_w$p.value, Sp_Po_Fr_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_Fr_p <- tibble(Country1, Country2, wilcox_Fr)
wilcox_Fr_p$holm_adjusted_w_Fr <- p.adjust(wilcox_Fr, method = "holm")
wilcox_Fr_p
```

### vegetables
```{r}
UK_Sw_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "UK", country2 = "Switzerland")
UK_Po_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "UK", country2 = "Poland")
UK_Sp_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "UK", country2 = "Spain")
Sw_Po_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "Switzerland", country2 = "Poland")
Sw_Sp_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "Switzerland", country2 = "Spain")
Sp_Po_V_w <- wilcox_pairwise(cat = "vegetables", country1 = "Spain", country2 = "Poland")
wilcox_V <- c(UK_Sw_V_w$p.value, UK_Po_V_w$p.value, UK_Sp_V_w$p.value, 
                Sw_Po_V_w$p.value, Sw_Sp_V_w$p.value, Sp_Po_V_w$p.value)
Country1 <- c("UK", "UK", "UK", "Switzerland", "Switzerland", "Spain")
Country2 <- c("Switzerland", "Poland", "Spain", "Poland", "Spain", "Poland")
wilcox_V_p <- tibble(Country1, Country2, wilcox_V)
wilcox_V_p$holm_adjusted_w_V <- p.adjust(wilcox_V, method = "holm")
wilcox_V_p
```

## Kruskal-Wallis testing without Switzerland
```{r}
DOMINO_MF_no_Sw <- filter(DOMINO_category, category == "meat_fish") %>%
  filter(!COUNTRY == "Switzerland")
KW_MF_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_MF_no_Sw)

DOMINO_BSB_no_Sw <- filter(DOMINO_category, category == "bread_sav_biscuits") %>%
  filter(!COUNTRY == "Switzerland")
KW_BSB_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_BSB_no_Sw)

DOMINO_C_no_Sw <- filter(DOMINO_category, category == "cereals") %>%
  filter(!COUNTRY == "Switzerland")
KW_C_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_C_no_Sw)

DOMINO_PRP_no_Sw <- filter(DOMINO_category, category == "potatoes_rice_pasta") %>%
  filter(!COUNTRY == "Switzerland")
KW_PRP_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_PRP_no_Sw)

DOMINO_DF_no_Sw <- filter(DOMINO_category, category == "dairy_fats") %>%
  filter(!COUNTRY == "Switzerland")
KW_DF_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_DF_no_Sw)

DOMINO_OBV_no_Sw <- filter(DOMINO_category, category == "on_bread_veg") %>%
  filter(!COUNTRY == "Switzerland")
KW_OBV_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_OBV_no_Sw)

DOMINO_SwS_no_Sw <- filter(DOMINO_category, category == "sweets_snacks") %>%
  filter(!COUNTRY == "Switzerland")
KW_SwS_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_SwS_no_Sw)

DOMINO_SSS_no_Sw <- filter(DOMINO_category, category == "soups_sauces_spreads") %>%
  filter(!COUNTRY == "Switzerland")
KW_SSS_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_SSS_no_Sw)

DOMINO_D_no_Sw <- filter(DOMINO_category, category == "drinks") %>%
  filter(!COUNTRY == "Switzerland")
KW_D_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_D_no_Sw)

DOMINO_Fr_no_Sw <- filter(DOMINO_category, category == "fruits") %>%
  filter(!COUNTRY == "Switzerland")
KW_Fr_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_Fr_no_Sw)

DOMINO_V_no_Sw <- filter(DOMINO_category, category == "vegetables") %>%
  filter(!COUNTRY == "Switzerland")
KW_V_no_Sw <- kruskal.test(FFQ ~ COUNTRY, data = DOMINO_V_no_Sw)

KW_p_no_Sw <- c(KW_MF_no_Sw$p.value, KW_BSB_no_Sw$p.value, KW_C_no_Sw$p.value, 
                KW_PRP_no_Sw$p.value, KW_DF_no_Sw$p.value, KW_OBV_no_Sw$p.value, 
                KW_SwS_no_Sw$p.value, KW_SSS_no_Sw$p.value, KW_D_no_Sw$p.value, 
                KW_Fr_no_Sw$p.value, KW_V_no_Sw$p.value)
KW_p_category_no_Sw <- tibble(KW_p_category, KW_p_no_Sw)
KW_p_category_no_Sw 
```

## Pairwise comparisons without Switzerland
### bread and savoury biscuits
```{r}
wilcox_pairwise <- function(cat, country1, country2) {
  DOMINO_category_no_Sw <- filter(DOMINO_category, category == cat) %>%
    filter(!COUNTRY == "Switzerland")
  DOMINO_category_no_Sw <- filter(DOMINO_category_no_Sw, COUNTRY == country1 | COUNTRY == country2)
  wilcox.test(FFQ ~ COUNTRY, data = DOMINO_category_no_Sw)
}


UK_Po_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "UK", country2 = "Poland")
UK_Sp_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "UK", country2 = "Spain")
Sp_Po_BSB_w <- wilcox_pairwise(cat = "bread_sav_biscuits", country1 = "Spain", country2 = "Poland")
wilcox_BSB_no_Sw <- c(UK_Po_BSB_w$p.value, UK_Sp_BSB_w$p.value,Sp_Po_BSB_w$p.value)
Country1 <- c("UK", "UK", "Spain")
Country2 <- c("Poland", "Spain", "Poland")
wilcox_BSB_p_no_Sw <- tibble(Country1, Country2, wilcox_BSB_no_Sw)
wilcox_BSB_p_no_Sw$holm_adjusted_w_BSB_no_Sw <- p.adjust(wilcox_BSB_no_Sw, method = "holm")
wilcox_BSB_p_no_Sw <- mutate(wilcox_BSB_p_no_Sw, "sig at p = 0.05?" = case_when(
                                                           holm_adjusted_w_BSB_no_Sw < 0.05 ~
                                                          "significant",  
                                                          holm_adjusted_w_BSB_no_Sw
                                                           > 0.05 ~ "not significant"))
wilcox_BSB_p_no_Sw
```

### cereals
```{r}
UK_Po_C_w <- wilcox_pairwise(cat = "cereals", country1 = "UK", country2 = "Poland")
UK_Sp_C_w <- wilcox_pairwise(cat = "cereals", country1 = "UK", country2 = "Spain")
Sp_Po_C_w <- wilcox_pairwise(cat = "cereals", country1 = "Spain", country2 = "Poland")
wilcox_C_no_Sw <- c(UK_Po_C_w$p.value, UK_Sp_C_w$p.value, Sp_Po_C_w$p.value)
Country1 <- c("UK", "UK", "Spain")
Country2 <- c("Poland", "Spain", "Poland")
wilcox_C_p_no_Sw <- tibble(Country1, Country2, wilcox_C_no_Sw)
wilcox_C_p_no_Sw$holm_adjusted_w_C_no_Sw <- p.adjust(wilcox_C_no_Sw, method = "holm")
wilcox_C_p_no_Sw <- mutate(wilcox_C_p_no_Sw, "sig at p = 0.05?" = case_when(
                                                           holm_adjusted_w_C_no_Sw < 0.05 ~
                                                          "significant",  
                                                          holm_adjusted_w_C_no_Sw
                                                           > 0.05 ~ "not significant"))
wilcox_C_p_no_Sw
```

### potatoes, rice and pasta
```{r}
UK_Po_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "UK", country2 = "Poland")
UK_Sp_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "UK", country2 = "Spain")
Sp_Po_PRP_w <- wilcox_pairwise(cat = "potatoes_rice_pasta", country1 = "Spain", country2 = "Poland")
wilcox_PRP_no_Sw <- c(UK_Po_PRP_w$p.value, UK_Sp_PRP_w$p.value, Sp_Po_PRP_w$p.value)
Country1 <- c("UK", "UK", "Spain")
Country2 <- c("Poland", "Spain", "Poland")
wilcox_PRP_p_no_Sw <- tibble(Country1, Country2, wilcox_PRP_no_Sw)
wilcox_PRP_p_no_Sw$holm_adjusted_w_PRP_no_Sw <- p.adjust(wilcox_PRP_no_Sw, method = "holm")
wilcox_PRP_p_no_Sw <- mutate(wilcox_PRP_p_no_Sw, "sig at p = 0.05?" = case_when(
                                                           holm_adjusted_w_PRP_no_Sw < 0.05 ~
                                                          "significant",  
                                                          holm_adjusted_w_PRP_no_Sw
                                                           > 0.05 ~ "not significant"))
wilcox_PRP_p_no_Sw
```

### dairy and fats
```{r}
UK_Po_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "UK", country2 = "Poland")
UK_Sp_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "UK", country2 = "Spain")
Sp_Po_DF_w <- wilcox_pairwise(cat = "dairy_fats", country1 = "Spain", country2 = "Poland")
wilcox_DF_no_Sw <- c(UK_Po_DF_w$p.value, UK_Sp_DF_w$p.value, Sp_Po_DF_w$p.value)
Country1 <- c("UK", "UK", "Spain")
Country2 <- c("Poland", "Spain", "Poland")
wilcox_DF_p_no_Sw <- tibble(Country1, Country2, wilcox_DF_no_Sw)
wilcox_DF_p_no_Sw$holm_adjusted_w_DF_no_Sw <- p.adjust(wilcox_DF_no_Sw, method = "holm")
wilcox_DF_p_no_Sw <- mutate(wilcox_DF_p_no_Sw, "sig at p = 0.05?" = case_when(
                                                           holm_adjusted_w_DF_no_Sw < 0.05 ~
                                                          "significant",  
                                                          holm_adjusted_w_DF_no_Sw
                                                           > 0.05 ~ "not significant"))
wilcox_DF_p_no_Sw
```

### sweets and snacks
```{r}
UK_Po_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "UK", country2 = "Poland")
UK_Sp_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "UK", country2 = "Spain")
Sp_Po_SwS_w <- wilcox_pairwise(cat = "sweets_snacks", country1 = "Spain", country2 = "Poland")
wilcox_SwS_no_Sw <- c(UK_Po_SwS_w$p.value, UK_Sp_SwS_w$p.value, Sp_Po_SwS_w$p.value)
Country1 <- c("UK", "UK", "Spain")
Country2 <- c("Poland", "Spain", "Poland")
wilcox_SwS_p_no_Sw <- tibble(Country1, Country2, wilcox_SwS_no_Sw)
wilcox_SwS_p_no_Sw$holm_adjusted_w_SwS_no_Sw <- p.adjust(wilcox_SwS_no_Sw, method = "holm")
wilcox_SwS_p_no_Sw <- mutate(wilcox_SwS_p_no_Sw, "sig at p = 0.05?" = case_when(
                                                           holm_adjusted_w_SwS_no_Sw < 0.05 ~
                                                          "significant",  
                                                          holm_adjusted_w_SwS_no_Sw
                                                           > 0.05 ~ "not significant"))
wilcox_SwS_p_no_Sw
```

### soups, sauces and spreads
```{r}
UK_Po_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "UK", country2 = "Poland")
UK_Sp_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "UK", country2 = "Spain")
Sp_Po_SSS_w <- wilcox_pairwise(cat = "soups_sauces_spreads", country1 = "Spain", country2 = "Poland")
wilcox_SSS_no_Sw <- c(UK_Po_SSS_w$p.value, UK_Sp_SSS_w$p.value, Sp_Po_SSS_w$p.value)
Country1 <- c("UK", "UK", "Spain")
Country2 <- c("Poland", "Spain", "Poland")
wilcox_SSS_p_no_Sw <- tibble(Country1, Country2, wilcox_SSS_no_Sw)
wilcox_SSS_p_no_Sw$holm_adjusted_w_SSS_no_Sw <- p.adjust(wilcox_SSS_no_Sw, method = "holm")
wilcox_SSS_p_no_Sw <- mutate(wilcox_SSS_p_no_Sw, "sig at p = 0.05?" = case_when(
                                                           holm_adjusted_w_SSS_no_Sw < 0.05 ~
                                                          "significant",  
                                                          holm_adjusted_w_SSS_no_Sw
                                                           > 0.05 ~ "not significant"))
wilcox_SSS_p_no_Sw
```

### drinks
```{r}
UK_Po_D_w <- wilcox_pairwise(cat = "drinks", country1 = "UK", country2 = "Poland")
UK_Sp_D_w <- wilcox_pairwise(cat = "drinks", country1 = "UK", country2 = "Spain")
Sp_Po_D_w <- wilcox_pairwise(cat = "drinks", country1 = "Spain", country2 = "Poland")
wilcox_D_no_Sw <- c(UK_Po_D_w$p.value, UK_Sp_D_w$p.value, Sp_Po_D_w$p.value)
Country1 <- c("UK", "UK", "Spain")
Country2 <- c("Poland", "Spain", "Poland")
wilcox_D_p_no_Sw <- tibble(Country1, Country2, wilcox_D_no_Sw)
wilcox_D_p_no_Sw$holm_adjusted_w_D_no_Sw <- p.adjust(wilcox_D_no_Sw, method = "holm")
wilcox_D_p_no_Sw <- mutate(wilcox_D_p_no_Sw, "sig at p = 0.05?" = case_when(
                                                           holm_adjusted_w_D_no_Sw < 0.05 ~
                                                          "significant",  
                                                          holm_adjusted_w_D_no_Sw
                                                           > 0.05 ~ "not significant"))
wilcox_D_p_no_Sw
```

### fruits
```{r}
UK_Po_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "UK", country2 = "Poland")
UK_Sp_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "UK", country2 = "Spain")
Sp_Po_Fr_w <- wilcox_pairwise(cat = "fruits", country1 = "Spain", country2 = "Poland")
wilcox_Fr_no_Sw <- c(UK_Po_Fr_w$p.value, UK_Sp_Fr_w$p.value, Sp_Po_Fr_w$p.value)
Country1 <- c("UK", "UK", "Spain")
Country2 <- c("Poland", "Spain", "Poland")
wilcox_Fr_p_no_Sw <- tibble(Country1, Country2, wilcox_Fr_no_Sw)
wilcox_Fr_p_no_Sw$holm_adjusted_w_Fr_no_Sw <- p.adjust(wilcox_Fr_no_Sw, method = "holm")
wilcox_Fr_p_no_Sw <- mutate(wilcox_Fr_p_no_Sw, "sig at p = 0.05?" = case_when(
                                                           holm_adjusted_w_Fr_no_Sw < 0.05 ~
                                                          "significant",  
                                                          holm_adjusted_w_Fr_no_Sw
                                                           > 0.05 ~ "not significant"))
wilcox_Fr_p_no_Sw
```

#### PCA
#### For all food groups excluding trout (28)
```{r,eval = FALSE}
library(devtools)
library(ggbiplot)
DOMINO_category_wide <- mutate(DOMINO, meat_fish = rowSums(DOMINO[2:28])) %>%
  mutate(bread_sav_biscuits = rowSums(DOMINO[29:35])) %>%
  mutate(cereals = rowSums(DOMINO[36:39])) %>%
  mutate(potatoes_rice_pasta = rowSums(DOMINO[40:57])) %>%
  mutate(dairy_fats = rowSums(DOMINO[58:79])) %>%
  mutate(on_bread_veg = rowSums(DOMINO[80:88])) %>%
  mutate(sweets_snacks = rowSums(DOMINO[89:115])) %>%
  mutate(soups_sauces_spreads = rowSums(DOMINO[116:128])) %>%
  mutate(drinks = rowSums(DOMINO[129:146])) %>%
  mutate(fruits = rowSums(DOMINO[147:163])) %>%
  mutate(vegetables = rowSums(DOMINO[164:196]))
DOMINO_category_wide <- DOMINO_category_wide[,-c(2:196)]

DOMINO_category_sd <- purrr::map_dbl(
  .x = DOMINO_category_wide, 
  .f = sd
) 

DOMINO_cat.pca <- prcomp((select(DOMINO_category_wide, !"COUNTRY")), center = TRUE, scale. = TRUE)
summary(DOMINO_cat.pca)
ggbiplot(DOMINO_cat.pca, ellipse = TRUE, obs.scale= 1, labels = rownames(DOMINO_category_wide), 
         groups = DOMINO_category_wide$COUNTRY)
```