---
title: "Meeting 1"
author: "Siobhan Lock"
date: "11/11/2021"
html_document:
  theme: sandstone
  toc: yes
  toc_float: yes
pdf_document:
  toc: yes
---

```{r setup, include=FALSE, echo = TRUE, error=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Meeting 1
### Data set-up
```{r,  include = FALSE, error=FALSE, warning=FALSE}
# Load packages and set wd
library(readr)
library(reshape2)
library(dplyr)
library(tidyverse)
library(stringr)
library(RColorBrewer)

setwd("D:/siobh/Documents/Uni/MSc/MET582/Coursework/Data Project")

### Import data and select prelimary variables (beef to olive)
DOMINO <- read_csv(file = "FFQ_for_MET582.csv", col_names = TRUE, na = "NA") %>%
  select(COUNTRY:OLIVES)
head(DOMINO)
table(DOMINO$COUNTRY)

```

```{r, include=TRUE, echo = FALSE, error=FALSE, warning=FALSE}
### Code missing values as NA and perform summary statistics
DOMINO[DOMINO == -9] <- NA
DOMINO[DOMINO == -8] <- NA

### Add participant IDs
DOMINO1 <- DOMINO
DOMINO1$ID <- seq.int(nrow(DOMINO1))
DOMINO1 <- DOMINO1 %>%
  select(COUNTRY, ID, BEEF:OLIVES)
summary(DOMINO1)
```

### Making a Summary Plot
```{r, fig.width=18, fig.height=8}
### Make summary plot showing food type by FFQ score, coloured by country.
DOMINO_grouped <- pivot_longer(data = DOMINO1, cols = BEEF:OLIVES, names_to = "FOOD", values_to = "FFQ")

food_plot <- ggplot (data = DOMINO_grouped, mapping = aes(x = FOOD, y = FFQ)) + 
  geom_col(aes(fill = COUNTRY)) +
  labs(title = "Cross-cultural Nutrional Differences") +
  xlab("Food Type") +
  ylab("FFQ Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  theme(legend.position="bottom") +
  scale_fill_brewer(palette="Spectral")
print(food_plot)
```

### Grouping food data by category.
```{r, include=TRUE, echo = FALSE, error=FALSE, warning=FALSE}
#### Group by major food groups.
MeatFish <- DOMINO1 %>%
  select(COUNTRY, ID, BEEF:TROUT) %>%
  pivot_longer(cols = BEEF:TROUT, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "MEAT & FISH")

Bread <- DOMINO1 %>%
  select(COUNTRY, ID, WHTBREADSL:RYEBRD) %>%
  pivot_longer(cols = WHTBREADSL:RYEBRD, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "BREAD & BISCUITS")

Cereals <- DOMINO1 %>%
  select(COUNTRY, ID, PORRIDGE:CEREALMUS) %>%
  pivot_longer(cols = PORRIDGE:CEREALMUS, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "CEREALS")

PotRicPas <- DOMINO1 %>%
  select(COUNTRY, ID, POTSPREP:BARLYGRO) %>%
  pivot_longer(cols = POTSPREP:BARLYGRO, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "POTATOES, RICE & PASTA")

DairyFats <- DOMINO1 %>%
  select(COUNTRY, ID, SICREAM:OTHSALAD) %>%
  pivot_longer(cols = SICREAM:OTHSALAD, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "DAIRY & FATS")

OnBreadVeg <- DOMINO1 %>%
  select(COUNTRY, ID, BUTTERMILK:VLOWSPREAD) %>%
  pivot_longer(cols = BUTTERMILK:VLOWSPREAD, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "TOPPING")

SweetSnacks <- DOMINO1 %>%
  select(COUNTRY, ID, SWBISCHO:JELLY) %>%
  pivot_longer(cols = SWBISCHO:JELLY, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "SWEETS & SNACKS")

SoupsSpreads <- DOMINO1 %>%
  select(COUNTRY, ID, VEGSOUP:COLDSOUP) %>%
  pivot_longer(cols = VEGSOUP:COLDSOUP, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "SOUPS & SPREADS")

Drinks <- DOMINO1 %>%
  select(COUNTRY, ID, TEA:TOMJUICE) %>%
  pivot_longer(cols = TEA:TOMJUICE, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "DRINKS")

Fruit <- DOMINO1 %>%
  select(COUNTRY, ID, APPLES:BLACKCURR) %>%
  pivot_longer(cols = APPLES:BLACKCURR, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "FRUIT")

Veg <- DOMINO1 %>%
  select(COUNTRY, ID, CARPUMP:OLIVES) %>%
  pivot_longer(cols = CARPUMP:OLIVES, names_to = "FOOD", values_to = "FFQ") %>%
  add_column(CATEGORY = "VEG")


# Create new table
DOMINO_bind <- rbind(MeatFish, Bread, Cereals, PotRicPas, DairyFats, OnBreadVeg, SweetSnacks, SoupsSpreads, Drinks, Fruit, Veg)

# Reorder
DOMINO_cat <- DOMINO_bind %>%
  select(ID, COUNTRY, CATEGORY, FOOD, FFQ)
head(DOMINO_cat)
```


### Category Summary Plots
```{r, include=TRUE, echo = FALSE, error=FALSE, warning=FALSE}
# More Plots
### Make summary plots showing food category by FFQ score, coloured by country.

cat_plot <- ggplot (data = DOMINO_cat, mapping = aes(x = CATEGORY, y = FFQ)) + 
  geom_col(aes(fill = COUNTRY)) +
  labs(title = "Cross-cultural Nutrional Differences by Food Category") +
  xlab("Food Category") +
  ylab("FFQ Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  scale_fill_brewer(palette="Spectral")
print(cat_plot)

cat_box <- ggplot(DOMINO_cat, aes(x=CATEGORY, y=FFQ, fill=COUNTRY)) +
  geom_boxplot(alpha=0.7) +
  stat_summary(fun=mean, geom="point", shape=20, size=14, color="red", fill="red") +
  labs(title = "Mean Cross-cultural Nutrional Differences by Food Category") +
  xlab("Food Category") +
  ylab("Mean FFQ Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  scale_fill_brewer(palette="Spectral")
print(cat_box)

cat_violin <- ggplot(DOMINO_cat, aes(x=CATEGORY, y=FFQ, fill=COUNTRY)) +
  geom_violin(trim = FALSE) +
  stat_summary(fun=mean, geom="point", shape=4, size=3,  colour="blue") +
  labs(title = "Mean Cross-cultural Nutrional Differences by Food Category") +
  xlab("Food Category") +
  ylab("Mean FFQ Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  scale_fill_brewer(palette="Spectral")
print(cat_violin)
```




# Meeting 2
Please note that although multiple Wilcoxen tests have been used to test for significant differences in the data (food category by country), no correction for multiple comparisons have been performed as of yet. These anslyses were completed as part of a lecture on hypothesis testing in our statistics module.

## A Start on Significance Testing 
### VEG
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON VEG CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "VEG")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "VEG")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (VEG) in UK and Spain")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON VEG CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "VEG")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (VEG) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON VEG CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (VEG) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```





### DAIRY & FAT
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON DAIRY & FATS CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "DAIRY & FATS")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "DAIRY & FATS")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (DAIRY & FATS) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON DAIRY & FATS CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "DAIRY & FATS")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (DAIRY & FATS) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON DAIRY & FATS CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (DAIRY & FATS) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```




### MEAT & FISH
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON MEAT & FISH CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "MEAT & FISH")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "MEAT & FISH")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (MEAT & FISH) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON MEAT & FISH CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "MEAT & FISH")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (MEAT & FISH) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON MEAT & FISH CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (MEAT & FISH) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```




### FRUIT
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON FRUIT CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "FRUIT")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "FRUIT")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (FRUIT) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON FRUIT CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "FRUIT")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (FRUIT) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON FRUIT CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (FRUIT) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```











### BREAD & BISCUITS
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON BREAD & BISCUITS CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "BREAD & BISCUITS")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "BREAD & BISCUITS")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (BREAD & BISCUITS) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON BREAD & BISCUITS CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "BREAD & BISCUITS")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (BREAD & BISCUITS) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON BREAD & BISCUITS CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (BREAD & BISCUITS) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```

















### CEREALS
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON CEREALS CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "CEREALS")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "CEREALS")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (CEREALS) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON CEREALS CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "CEREALS")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (CEREALS) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON CEREALS CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (CEREALS) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```





### DRINKS
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON DRINKS CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "DRINKS")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "DRINKS")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (DRINKS) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON DRINKS CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "DRINKS")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (DRINKS) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON DRINKS CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (DRINKS) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```



















### POTATOES, RICE & PASTA
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON POTATOES, RICE & PASTA CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "POTATOES, RICE & PASTA")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "POTATOES, RICE & PASTA")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (POTATOES, RICE & PASTA) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON POTATOES, RICE & PASTA CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "POTATOES, RICE & PASTA")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (POTATOES, RICE & PASTA) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON POTATOES, RICE & PASTA CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (POTATOES, RICE & PASTA) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```


















### SOUPS & SPREADS
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON SOUPS & SPREADS CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "SOUPS & SPREADS")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "SOUPS & SPREADS")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (SOUPS & SPREADS) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON SOUPS & SPREADS CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "SOUPS & SPREADS")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (SOUPS & SPREADS) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON SOUPS & SPREADS CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (SOUPS & SPREADS) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```


















### SWEETS & SNACKS
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON SWEETS & SNACKS CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "SWEETS & SNACKS")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "SWEETS & SNACKS")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (SWEETS & SNACKS) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON SWEETS & SNACKS CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "SWEETS & SNACKS")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (SWEETS & SNACKS) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON SWEETS & SNACKS CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (SWEETS & SNACKS) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```


















### TOPPING
```{r}
library(sm)
sm.density(DOMINO_cat$FFQ, xlab = "FFQ Score")
shapiro.test(DOMINO_cat$FFQ[0:5000])

library(nortest)
ad.test(DOMINO_cat$FFQ)

# Significant result indicates data comes from a non-normal distribution (which matches the shape of the density plot)

# tESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON TOPPING CONSUMPTION
UK_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "UK" & CATEGORY == "TOPPING")
SPAIN_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Spain" & CATEGORY == "TOPPING")
test_DOMINO <- rbind(UK_DOMINO, SPAIN_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (TOPPING) in UK and Spain ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON TOPPING CONSUMPTION
POL_DOMINO <- DOMINO_cat %>%
  filter(COUNTRY == "Poland" & CATEGORY == "TOPPING")
test_DOMINO <- rbind(UK_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (TOPPING) in UK and Poland ")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)

# TESTING WHETHER SPAIN AND UK SIGNIFICANTLY DIFFER ON TOPPING CONSUMPTION
test_DOMINO <- rbind(SPAIN_DOMINO, POL_DOMINO)
wilcox.test(test_DOMINO$FFQ ~ test_DOMINO$COUNTRY)

ggplot(test_DOMINO, aes(COUNTRY,FFQ, color=COUNTRY))+stat_summary(fun.y=mean,geom="point")+
    stat_summary(fun.data=mean_cl_normal, geom="errorbar")+ylab("Mean FFQ Score")+xlab("Country")+ggtitle("FFQ Scores (TOPPING) in Spain and Poland")+  theme(legend.position = "none")

# Delineate
table(test_DOMINO$COUNTRY)
```

## Attempting a PCA Analysis:

```{r}
library(tibble)
library(devtools)
library(ggbiplot)

# PCA with all food types
DOMINO_food <- DOMINO_grouped %>%
  na.omit() %>%
  pivot_wider(id_cols = COUNTRY:ID, names_from = FOOD, values_from = FFQ) %>%
  select(BEEF:OLIVES)
 
DOMINO_pca <- prcomp(DOMINO_food, center = TRUE, scale. = TRUE)

summary(DOMINO_pca)
DOMINO_countries <- c(rep("UK", 10), rep("Switzerland", 2), rep("Poland", 18), rep("Spain", 30))
DOMINO_bp <- ggbiplot(DOMINO_pca, ellipse = TRUE,  groups = DOMINO_countries)
print(DOMINO_bp)

# Lots of labels so try with categories. 
DOMINO_wide <- DOMINO_cat %>%
  na.omit() %>%
  select(-FOOD) %>%
  pivot_wider(id_cols = ID:COUNTRY, names_from = CATEGORY, values_from = FFQ, values_fn = sum) %>%
  select(3:13)

DOMINO_cat_pca <- prcomp(DOMINO_wide, center = TRUE, scale. = TRUE)
summary(DOMINO_cat_pca)
DOMINO_cat_bp <- ggbiplot(DOMINO_cat_pca, ellipse = TRUE,  groups = DOMINO_countries)
print(DOMINO_cat_bp)
```



## Summary of food category Information.
```{r}
summary(DOMINO_wide)
mean_cat_plot <- ggplot (data = DOMINO_cat, mapping = aes(x = CATEGORY, y = FFQ)) + 
  geom_col(aes(fill = COUNTRY), position = "dodge", stat = "mean") +
  labs(title = "Cross-cultural Nutrional Differences by Food Category") +
  xlab("Food Category") +
  ylab("Mean FFQ Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
  scale_fill_brewer(palette="Spectral")
print(mean_cat_plot)
```

